\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2022-04-04 lin-vs-cen.tex
Last minute airplane stuff

\begin{lem}
  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.\todo{Use a specific value of $d$}.
  Then there exists $S\subseteq V(G)$ such that
  \begin{compactenum}[(i)]
    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|\ge 2$; and
    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
  \end{compactenum}
\end{lem}

@@

  % To begin, define a supergraph $H^+\supseteq H$ that contains an edge $\alpha\mu$ if there exists at least one $\nu\in\VE(G_r(\mu))\cap X$ with $\phi(\nu)=\alpha$.

@@

  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and $\nu_1,\ldots,\nu_r$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$ and $\phi(\nu_i)=\alpha$, for each $i\in\{1,\ldots,t\}$.  The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.  Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears before any elements of $X_{\nu_i}$ in our random permutation of $Y$'' then the events $U_1,\ldots,U_k$ are mutually independent.  Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.  For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/?$.  Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/?)$ random variable.  Therefore,
  \begin{align*}
    \Pr\left(\deg_H(\alpha)< 100r^2\right)
    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
    & \le \Pr(\binomial(t,1/?)\le 100r^2) \\
    & \le (2t)^{100r^2}(1/?)^{r^2}(1-1/?)^{t-100r^2} \\
    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/?)) \\
    & = \exp(ar^2 - bt)
  \end{align*}
  for fixed positive $a,b>0$.
  \todo[inline]{Work the preceding calculation out}
  We are now ready for an application of \cref{weighted_lovasz}.  For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $Z_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.  The event $X_\alpha$ is mutually independent of $\Gamma_\alpha$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le ?|Z_\alpha|$.

@@

  For each $\alpha\in X$, define $w(E_\alpha)=c/2r^2$ and observe that
  \begin{align*}
    p(1-p)^{|\Gamma_\alpha|}
    & \le  p(1-p)^{?|Z_\alpha|} = \exp(\log p + ?|Z_\alpha|\log(1-p))
  \end{align*}

  % (Think of $\nu$ as belonging to the first box that covers it.) Now, for each $\alpha\in X$, $\deg_H(\alpha)\ge d-1$ since, otherwise the first round would have been successful for $\alpha$.  Straightforward counting shows that, for each $\mu\in Y$,  $\deg_H(\mu)\le |\VE(G_r(\mu))|\le 2(6r^2+6r+1)=:C$.

  If $d$ and $r$ are chosen so that $(d-1)/C\ge 2$ then \cref{d_hall} guarantees that we can find a set $Q_2$ so that $Q_1\cup Q_2$ satisfies (\ref{hits_both}) and such that $|B_r(\mu)\cap Q_2|\le 1$ for each $u\in Q_1$, which comes very close to satisfying (\ref{spread_out}).  Unfortunately, this approach is too naïve: it can produce triples (and even $4$-tuples) of objects in $Q$ are all contained in a small box and therefore do not satisfy (\ref{spread_out}).  We will overcome this difficulty with an application of the Lovász Local Lemma.

  Consider a (for now) random permutation $\mu_1,\ldots,\mu_{2|K|}$ of $Q_1$.  We remove any edge from $\alpha\mu_i$ from $H$ if
  \[
    \alpha\not\in \phi(\VE(G_r(\mu_i)\setminus(\bigcup_{j=1}^{i-1}\VE(G_{2r}(\mu_j))))
  \]
  contains no object $\nu$ with $\phi(\nu)=\alpha$.  (Note the extra factor of $2$ in the radii on the right hand side.)
  The effect of this is that each $\mu_j\in Q_1$ defines a box of radius $2r$ around $\mu_j$ that is forbidden after the $j$th step of the second round.  We will now use \cref{weighted_lovasz} to prove that, with some positive probability, each vertex

  find $2(r+1)^2$ elements $\mu\in S_1$ such that
  \begin{compactenum}[(a)]
    \item
  \end{compactenum}
\end{proof}

>>>>

\begin{lem}
  For every integer $r\ge 1$ there exists $d\in O(r^4)$ such that the following is true.
  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.\todo{Use a specific value of $d$}.
  Then there exists $S\subseteq V(G)$ such that
  \begin{compactenum}[(i)]
    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|\ge 2$; and
    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
  \end{compactenum}
\end{lem}

@@

@@

  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and $\nu_1,\ldots,\nu_r$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$ and $\phi(\nu_i)=\alpha$, for each $i\in\{1,\ldots,t\}$.  The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.  Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears before any elements of $X_{\nu_i}$ in our random permutation of $Y$'' then the events $U_1,\ldots,U_k$ are mutually independent.  Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.  For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/?$.  Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/?)$ random variable.  Therefore,
  \begin{align*}
    \Pr\left(\deg_H(\alpha)< 100r^2\right)
    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
    & \le \Pr(\binomial(t,1/?)\le 100r^2) \\
    & \le (2t)^{100r^2}(1/?)^{r^2}(1-1/?)^{t-100r^2} \\
    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/?)) \\
    & = \exp(ar^2 - bt)
  \end{align*}
  for fixed positive $a,b>0$.
  \todo[inline]{Work the preceding calculation out}
  We are now ready for an application of \cref{weighted_lovasz}.  For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $Z_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.  The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le ?|Z_\alpha|$.

@@

  For each $\alpha\in X$, define $w(E_\alpha)=c/2r^2$ and observe that
  \begin{align*}
    p(1-p)^{|\Gamma_\alpha|}
    & \ge  p(1-p)^{?|Z_\alpha|} = \exp(\log c/2r^2 + ?|Z_\alpha|\log(1-c/2r^2))
  \end{align*}
  \todo[inline]{Sort out these calculations}

  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 100r^2$ for each $\alpha\in X$.  On the other hand, $\deg_{H}(y)\le 8r^2$ for each $y\in Y$.  Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 100|A|/8 \ge 2|A|$.  Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$. Each edge $\alpha\mu$ of $M$ corresponds to some $\nu\in X_\alpha\cap\VE(G_r(\mu))$, which we place in $Q_2$.

  Now it is straightforward to verify that $Q:=Q_1\cup Q_2$ satisifies (\ref{hits_both}).  Furthermore, for each $\mu\in Q_1$, $B_r(\mu)$ contains no object in $Q_1\setminus\{\mu\}$ and contains at most one element of $Q_2$.  Similarly, for each $\mu\in Q_2$, $B_r(\mu)$ contains no element of $Q_2\setminus\{\mu\}$ and contains at most one element of $Q_1$.  Therefore $Q$ satisfies (\ref{spread_out}).
\end{proof}
